---
title: "PA1_template"
author: "Michael Rith"
date: "July 12, 2016"
output: 
        md_document:
        variant: markdown_github
---
```{r}
setwd("C:/Users/mrith/Documents/ReproducableResearch")
#Open Data Set
data <- read.csv("activity.csv")
#install lattice package
library(lattice)
```
Create Histogram with Mean and Median values for number of steps
```{r}
#Sum steps by day, create Histogram, and calculate mean and median.
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
rmean <- mean(steps_by_day$steps)
rmedian <- median(steps_by_day$steps)
```

The Mean is 10766 and the Median is 10765

#What is the average daily activity?

```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, 
     type="l", xlab="Interval",
     ylab="Number of Steps",
     main="Average # of Steps per Day by Interval")
```
Remove all missing variables and imputed zeros
```{r}
removeNA <- sum(!complete.cases(data))
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
```
Total Step recount and create a Histogram
```{r}
steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")

#Create Histogram graph to show difference. 
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="orange", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("green", "orange"), lwd=10)
```
Calculate new mean and median for imputed data.

```{r}

rmean.i <- mean(steps_by_day_i$steps)
rmedian.i <- median(steps_by_day_i$steps)

```

Calculate difference between imputed and non-imputed data.
```{r}

mean_diff <- rmean.i - rmean
med_diff <- rmedian.i - rmedian

```
Calculate total difference.

```{r}

total_diff <- sum(steps_by_day_i$steps) - sum(steps_by_day$steps)

```
* The imputed data mean is 1.059 × $10^{4}$
* The imputed data median is 1.0766 × $10^{4}$
* The difference between the non-imputed mean and imputed mean is -176.4949
* The difference between the non-imputed mean and imputed mean is 1.1887
* The difference between total number of steps between imputed and non-imputed data is 7.5363 × $10^{4}$.Therefore, there were 7.5363 × $10^{4}$ more steps in the imputed data.

#Are there differences in activity patterns between weekdays and weekends?
Want to create a plot to compare and contrast number of steps between the week and weekend. There is a higher peak earlier on weekdays, and more overall activity on weekends.
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)


xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, 
       main="Average Steps per Day by Interval",xlab="Interval", 
       ylab="Steps",layout=c(1,2), type="l")

```



You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

* The imputed data mean is `r mean_diff`

